---
title: "DATA 605 - Assignment 2"
author: "Joshua Sturm"
date: "02/05/2018"
output:
  pdf_document:
    keep_tex: yes
  html_document:
    highlight: pygments
    theme: cerulean
    code_folding: show
    toc: yes
    toc_float: yes
    smart: no
always_allow_html: yes
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = F, message = F)
```

# Problem Set 1

## 1. Show that $A^TA \neq AA^T$ in general. (Proof and demonstration.)
### Solution

Let $A = \begin{bmatrix}a & b\\ c & d\end{bmatrix}.$ Then $A^T = \begin{bmatrix}a & c\\ b & d\end{bmatrix}$.

The product of $AA^T$ will be: $\begin{bmatrix}a\times a + b\times b & a\times c + b\times d\\ c\times a + d\times b & c\times c + d \times d\end{bmatrix}$.

The product of $A^TA$ will be: $\begin{bmatrix}a\times a + c\times c & a\times b + c\times d\\ a\times b + c\times d & b\times b + d \times d\end{bmatrix}$.

Thus, we can see that matrix multiplication does not have the commutative property.

In the event that $A$ is not a square matrix, then the proof is trivial. Say $B$ is an $m\times n$ matrix, then $B^T$ will be an $n\times m$ matrix. $BB^T$ will thus be an $m\times m$ matrix, which is a square matrix.

## 2. For a special type of square matrix $A$, we get $A^TA = AA^T$. Under what conditions could this be true? (Hint: The identity matrix $I$ is an example of such a matrix.)
### Solution

Theorem: For any matrix $A$, $AA^T$ and $A^TA$ are symmetric matrices.\
The product of two symmetric matrices is symmetric only if both matrices commute, i.e. $AA^T = A^TA$.\
(https://en.wikipedia.org/wiki/Symmetric_matrix#Properties)\
This is only possible when the matrix is invertible.

# Problem Set 3
Write an R function to factorize a square matrix A into LU or LDU, whichever you prefer. You don’t have to worry about permuting rows of A and you can assume that A is less than 5x5, if you need to hard-code any variables in your code

```{r ps3}

# Assume the user inputs a square matrix.
  
factorize <- function(mat){
  L <- diag(nrow(mat))    # Create lower diagonal matrix of the same size
  U <- mat

  # If the leading entry [1,1] is not > 1, swap it with a row that has a pivot
  if (U[1,1] == 0){
    for (i in U[2,1]:U[nrow,1]){
      if (U[i,1] != 0){
        U[1,] <- U[i,]
      }
    }
  }
  # Loop through n-1 columns, n rows, get the multipler,
  # Multiply to get the Upper Triangular matrix, and plug
  # the multiplier into the Lower Triangular Matrix
  for (k in 1:(nrow(U)-1)){
    for (j in (k+1):(nrow(U))){
      if (U[j,k] != 0){
        mult <- U[j,k] / U[k,k]
        U[j,] <- U[j,] - (mult * U[k,])
        L[j,k] <- mult
      }
    }
  }
  
  factored.matrix <- list(L, U)
  
  return(factored.matrix)
}
```