---
title: "DATA 605 - Discussion 4 response"
author: "Joshua Sturm"
date: "02/22/2018"
output:
  pdf_document:
    keep_tex: yes
  html_document:
    highlight: pygments
    theme: cerulean
    code_folding: show
    toc: yes
    toc_float: yes
    smart: no
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = F, message = F)
```

Page 460, Exercise C25  

## Define the linear transformation

$T: \mathbb{C}^3 \to \mathbb{C}^2, \hspace{2mm} T\left(\begin{bmatrix}x_1\\x_2\\x_3\end{bmatrix}\right) = \begin{bmatrix}2x_1 - x_2 + 5x_3\\ -4x_1 + 2x_2 - 10x_3\end{bmatrix}$.

Find a basis for the kernel of $T, \ \mathcal{K}(T)$. Is $T$ injective?

Set $T(x) = 0$, and then row reduce.

\[
\left[\begin{array}{ccc|c}
  2 & -1 & 5 & 0 \\
  -4 & 2 & -10 & 0
\end{array} \right]
\]

```{r}
library(pracma)

A <- matrix(c(2, -1, 5, 0,
              -4, 2, -10, 0),
             nrow = 2, ncol = 4, byrow = T)

a <- rref(A)
a
```

Which gives us the system of equations:

\begin{align*}
x_1 - \frac{1}{2}x_2 + \frac{5}{2}x_3 &= 0 \\ 
0 &= 0
\end{align*}

So $x_1 = \frac{1}{2}x_2 - \frac{5}{2}x_3$, with $x_2$ and $x_3$ being free variables.

Thus, the basis set for the null space is $\left\{\begin{bmatrix}\frac{1}{2}\\1\\0\end{bmatrix}, \begin{bmatrix}\frac{-5}{2}\\0\\1\end{bmatrix}\right\}$

Theorem KILT says $T$ is injective iff $\mathcal{K}(T)=\{0\}$.

Since the kernel is not trivial, i.e. $\mathcal{K}(T)\neq\{0\}, \ T$ is not injective.