---
title: "DATA 605 - Assignment 11"
author: "Joshua Sturm"
date: "April 22, 2018"
output:
  html_document:
    highlight: textmate
    theme: sandstone
    code_folding: show
    toc: yes
    toc_float: yes
    smart: yes
  github_document:
  pdf_document:
    keep_tex: yes
always_allow_html: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = F, message = F, collapse = T, cache = T)
```

Using the “cars” dataset in R, build a linear model for stopping distance as a function of speed and replicate the analysis of your textbook chapter 3 (visualization, quality evaluation of the model, and residual analysis.)

```{r load-libraries}
library(tidyverse)
library(gridExtra)
```

```{r load-dataset}
data <- cars
glimpse(data)
```
The dataset contains 2 variables, and 50 cases.

```{r build-model}
model <- lm(formula = dist ~ speed,
            data = data)
summary(model)
```
The model has an adjusted r-squared of `r summary(model)$adj.r.squared`, and a p-value of `r summary(model)$coefficients[2,4]`.

```{r predict-plots}

rp1 <- ggplot(model, aes(.fitted, .resid)) +
  geom_point() +
  geom_hline(yintercept = 0) +
  geom_smooth(se = FALSE) +
  labs(title = "Residuals vs Fitted")

rp2 <- ggplot(model, aes(.fitted, .stdresid)) +
  geom_point() +
  geom_hline(yintercept = 0) +
  geom_smooth(se = FALSE)

rp3 <- ggplot(model) +
  stat_qq(aes(sample = .stdresid)) +
  geom_abline()

rp4 <- ggplot(model, aes(.fitted, sqrt(abs(.stdresid)))) +
  geom_point() +
  geom_smooth(se = FALSE) +
  labs(title = "Scale-Location")

rp5 <- ggplot(model, aes(seq_along(.cooksd), .cooksd)) +
  geom_col() +
  ylim(0, 0.0075) +
  labs(title = "Cook's Distance")

rp6 <- ggplot(model, aes(.hat, .stdresid)) +
  geom_point(aes(size = .cooksd)) +
  geom_smooth(se = FALSE, size = 0.5) +
  labs(title = "Residuals vs Leverage")

rp7 <- ggplot(model, aes(.hat, .cooksd)) +
  geom_vline(xintercept = 0, colour = NA) +
  geom_abline(slope = seq(0, 3, by = 0.5), colour = "white") +
  geom_smooth(se = FALSE) +
  geom_point() +
  labs(title = "Cook's distance vs Leverage")

rp8 <- ggplot(model, aes(.resid)) +
  geom_histogram(bins = 7, color="darkblue", fill="steelblue")

grid.arrange(rp1, rp2, rp3, rp4, rp5, rp6, rp7, rp8, ncol = 2)
```

The residuals appear to be nearly normal, as can be seen in the plots `fitted vs. residuals`, `Q-Q Plot`, and the residual histogram. Taken together with the information from the model summary, we can conclude that this model is sufficiently capable of making predictions on this dataset.
