---
title: "DATA 605 - Assignment 3"
author: "Joshua Sturm"
date: "02/18/2018"
output:
  pdf_document:
    keep_tex: yes
  html_document:
    highlight: pygments
    theme: cerulean
    code_folding: show
    toc: yes
    toc_float: yes
    smart: no
always_allow_html: yes
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = F, message = F, collapse = T)
```

```{r load-packages}
library(pracma)
```
# Problem Set 1

## 1. What is the rank of matrix $A$?

$$A = \begin{bmatrix} 1 & 2 & 3 & 4\\
                     -1 & 0 & 1 & 3\\
                     0 & 1 & -2 & 1 \\
                     5 & 4 & -2 & -3
     \end{bmatrix}$$
     
```{r}
A <- matrix(c(1, 2, 3, 4,
             -1, 0, 1, 3,
             0, 1, -2, 1,
             5, 4, -2, -3),
           nrow = 4, ncol = 4, byrow = T)

rref(A)
```
There are 4 pivot columns in the reduced row echelon form of the matrix, so it has a rank of 4.

## 2. Given an $m\times n$ matrix where $m>n$, what can be the maximum rank? The minimum rank, assuming that the matrix is non-zero?

Since the rank is the number of linearly independent columns, the maximum rank can't be larger than the number of columns, i.e. $n$.

If the matrix is non-zero, the minimum rank would be 1.

## 3. What is the rank of matrix $B$?

$$B = \begin{bmatrix} 1 & 2 & 1\\
                     3 & 6 & 3\\
                     2 & 4 & 2
     \end{bmatrix}$$
     
Using the same process as question 1:

```{r}
B <- matrix(c(1, 2, 1,
              3, 6, 3,
              2, 4, 2),
            nrow = 3, ncol = 3, byrow = T)

rref(B)
```

Since the reduced row echelon form of $B$ only has one pivot column, its rank is 1.

# Problem Set 2

## 1. Compute the eigenvalues and eigenvectors of the matrix $A$. You'll need to show your work. You'll need to write out the characteristic polynomial and show your solution.

$$A = \begin{bmatrix} 1 & 2 & 3\\
                     0 & 4 & 5\\
                     0 & 0 & 6
     \end{bmatrix}$$
     
$$|A - \lambda I | = \left | \begin{bmatrix} 1 & 2 & 3\\ 0 & 4 & 5\\0 & 0 & 6\end{bmatrix} - \lambda \begin{bmatrix}1 & 0 & 0 \\ 0 & 1 & 0\\ 0 & 0 & 1 \end{bmatrix} \right |$$
                     
$= \left|\begin{bmatrix}1-\lambda & 2 & 3\\ 0 & 4-\lambda & 5\\ 0 & 0 & 6-\lambda\end{bmatrix}\right|$    

$= (1-\lambda)(4-\lambda)(6-\lambda)$

$(1 - \lambda)\Big[24-10\lambda+\lambda^2\Big] = 24 -10\lambda + \lambda^2 -24\lambda +10\lambda^2 - \lambda^3 = -\lambda^3 + 11\lambda^2 -34\lambda + 24$

After factoring, we get $(\lambda - 1)(\lambda - 4)(\lambda - 6)$.

So the eigenvalues are 1, 4, 6.

For $\lambda = 1$:
     
```{r}
A <- matrix(c(1, 2, 3,
              0, 4, 5,
              0, 0, 6),
            nrow = 3, ncol = 3, byrow = T)

I <- diag(3)

rref(A-I)
```

$v_1 = \begin{bmatrix}1\\0\\0\\\end{bmatrix}$

For $\lambda = 4$:

```{r}
rref(A - 4*I)
```

$x_1 - \frac{2}{3}x_2 = 0$

$v_4 = \begin{bmatrix}\frac{2}{3}\\1\\0\end{bmatrix}$

For $\lambda = 6$:

```{r}
rref(A - 6*I)
```

$x_1 - 1.6x_3 = 0 \hspace{20mm} x_2 - 2.5x_3 = 0 \hspace{5mm} x_3\neq 0$.

$v_6 = \begin{bmatrix}1.6 \\ 2.5 \\ 1 \end{bmatrix}$