---
title: "DATA 605 - Discussion 1"
author: "Joshua Sturm"
date: "01/31/2018"
output:
  html_document:
    highlight: pygments
    theme: cerulean
    code_folding: show
    toc: yes
    toc_float: yes
    smart: no
  pdf_document:
    keep_tex: yes
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = F, message = F)
```

Page 43, problem C31.

Row-reduce the matrix without the aid of a calculator, indicating
the row operations you are using at each step using the notation of Definition RO.

I'll use Gauss-Jordan elimination.

$$\begin{bmatrix}
1 & 2 & -4 \\
-3 & -1 & -3 \\
-2 & 1 & -7
\end{bmatrix}$$

$$R_2 = R_2 + 3R_1, \ \ R_3 = R_3 + 2R_1
\begin{bmatrix}
1 & 2 & -4 \\
0 & 5 & -15 \\
0 & 5 & -15
\end{bmatrix}$$

$$R_2 = R_2\times \frac{1}{5} \
\begin{bmatrix}
1 & 2 & -4 \\
0 & 1 & -3 \\
0 & 5 & -15
\end{bmatrix}$$

$$R_3 = R_3 - 5R_2
\begin{bmatrix}
1 & 2 & -4 \\
0 & 1 & -3 \\
0 & 0 & 0
\end{bmatrix}$$

$$R_1 = R_1 -2R_2
\begin{bmatrix}
1 & 0 & 2 \\
0 & 1 & -3 \\
0 & 0 & 0
\end{bmatrix}$$

The matrix is now in reduced row echelon form.

To verify, we can perform the operations in R.

```{r ex31}
# Set up our matrix
a <- matrix(c(1,2,-4,-3,-1,-3,-2,1,-7), ncol=3, nrow=3, byrow=T)
a
```

I found this row-reducing function on github (https://gist.github.com/ZPears/0583aae73aa06d8abd9e).

```{r gj-formula}
convertToRREF <- function(matrix) {

  m <- nrow(matrix)
  n <- ncol(matrix)
  currCol <- 1
  nonZeroRowCount <- 0
  
  while ( (currCol < n+1) & (nonZeroRowCount+1 <= m) )  {

    if (sum(matrix[(nonZeroRowCount+1):m, currCol]) == 0) {
      
      currCol <- currCol + 1

    } else {
      
      rowIndex <- 0
    
      for (i in nonZeroRowCount+1:m) {
        
        if (matrix[i,currCol] != 0) {
          
          rowIndex <- i
          break
          
        }
        
      }
      
      nonZeroRowCount <- nonZeroRowCount + 1
      
      # switch rows
      
      row1 <- matrix[rowIndex,]
      row2 <- matrix[nonZeroRowCount,]
      matrix[rowIndex,] <- row2
      matrix[nonZeroRowCount,] <- row1
      
      # Use the second row operation (mult. by scalar) 
      # to convert the entry in row 
      # nonZeroRowCount and column currCol to a 1.
      
      matrix[nonZeroRowCount,] <- (1/matrix[nonZeroRowCount,currCol]) * matrix[nonZeroRowCount,]

      # Use the third row operation with row nonZeroRowCount
      # to convert every other entry of column j to zero.
      
      for (k in 1:m) {
        
        if ( (matrix[k, currCol] != 0) & (k != nonZeroRowCount) ) {
          
          # use row r (nonZeroRowCount) and row op 3 to make A[k, currCol] = 0
          
          scalar <- matrix[k, currCol] / matrix[nonZeroRowCount, currCol] 
          
          matrix[k, ] <- -1 * scalar * matrix[nonZeroRowCount, ] + matrix[k, ]

        }
        
      }
      
      # increment and repeat
      
      currCol <- currCol + 1
      
    }
        
  }
  
  return(matrix)
      
}
```

Finally, let's verify that we get the same answer.

```{r verify}
convertToRREF(a)
```