---
title: "DATA 605 - Assignment 7"
author: "Joshua Sturm"
date: "March 15, 2018"
output:
  pdf_document:
    keep_tex: yes
  html_document:
    highlight: pygments
    theme: cerulean
    code_folding: show
    toc: yes
    toc_float: yes
    smart: no
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = F, message = F, collapse = T)
```

# 1.
Let $X_1, X_2, . . . , X_n$ be $n$ mutually independent random variables, each of which is uniformly distributed on the integers from $1$ to $k$. Let $Y$ denote the minimum of the $X_i$’s. Find the distribution of $Y$.

## Solution
We first consider how many ways we can assign $Y$ to one of the $x_i$ variables.

Since there are $n$ variables, with $k$ values, there are $k^n$ total possibilities.

The number of ways we can get $Y = 1$ is the total number of possibilities less the number of possibilities that $Y$ is \textit{not} $1$. So, this can be done in $k^{n} - (k-1)^n$ ways. 

Similarly, the number of ways to get $Y = 2$ would be $k^{n} - (k-2)^{n} - [k^{n} - (k-1)^{n}] = (k-1)^{n} - (k-2)^{n}$.

The pattern we get is the recurrence relation $P(Y = m) = \frac{(k-m+1)^{n} - (k-m)^{n}}{k^{n}}$.

# 2.
Your organization owns a copier (future lawyers, etc.) or MRI (future doctors). This machine has a manufacturer’s expected lifetime of 10 years. This means that we expect one failure every ten years. (Include the probability statements and R Code for each part.)

## a.
What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as a geometric. (Hint: the probability is equivalent to not failing during the first 8 years.)

### Solution
Let $p$ be the probability that the machine fails, and $q = 1 - p$ the probability that it doesn't.

We're looking for the first failure (success) after 8 years, so this is a geometric distribution. Since we're only expecting one failure in 10 years, $p = 0.1, \ q = 0.9$.

$P(X > 8) = 1 - P(X \leq 8)$

$1 - (1 - q^{i+1}) = q^{i+1} = 0.9^{9} =$ `r round(0.9^9, 4)`.

$E[X] = \frac{1}{p} = \frac{1}{0.1} =$ `r 1/0.1`.

$Var(X) = \frac{1-p}{p^2} = \frac{0.9}{(0.1)^2} =$ `r 0.9/(0.1)^2`.

Standard deviation = $\sqrt{Var(X)} = \sqrt{90} \approx$ `r sqrt(90)`.

```{r part-a}
pdf <- pgeom(8, 0.1, lower.tail = F)

p <- 0.1
q <- 1 - p
ex <- p^-1
var <- q/p^2
sd <- sqrt(var)

cat(sprintf("\n %s = %f \n", 
            c("Probability", "Expected Value", "Variance", "Standard Deviation"),
            c(pdf, ex, var, sd))
    )
```

## b.
What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as an exponential. 

### Solution
$P(X > 8) = e^{-\lambda i}$, with $\lambda = 0.1$.

$P(X > 8) = e^{-0.8} =$ `r exp(-0.8)`.

$E[X] = \frac{1}{\lambda} =$ `r 1/0.1`.

$Var(X) = \frac{1}{\lambda^2} = \frac{1}{0.1^2} =$ `r 1/(0.1)^2`.

Standard Deviation $= \sqrt{Var(X)} = \sqrt{100} = 10$.

```{r part-b}
pdf <- pexp(8, 0.1, lower.tail = F)

l <- 0.1
ex <- 1/l
var <- 1/l^2
sd <- sqrt(var)

cat(sprintf("\n %s = %f \n", 
            c("Probability", "Expected Value", "Variance", "Standard Deviation"),
            c(pdf, ex, var, sd))
    )
```

## c.
What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as a binomial. (Hint: 0 success in 8 years)

### Solution
We're looking for 0 successes in 8 years. So, $P(0)$ with $n = 8$.

$P(0) = \binom{8}{0}(0.1)^0 \times 0.9{8-0} = (0.9)^8 =$ `r (0.9)^8`.

$E[X] = np = 8\cdot 0.1 = 0.8$.

$Var(X) = npq = 0.8\cdot 0.9 =$ `r 0.8*0.9`.

Standard Deviation $= \sqrt{Var(X)} = \sqrt{0.72} =$ `r sqrt(0.72)`.

```{r part-c}
pdf <- pbinom(0, 8, 0.1)

n <- 8
i <- 0
p <- 0.1
q <- 0.9
ex <- n*p
var <- n*p*q
sd <- sqrt(var)

cat(sprintf("\n %s = %f \n", 
            c("Probability", "Expected Value", "Variance", "Standard Deviation"),
            c(pdf, ex, var, sd))
    )
```

## d.
What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as a Poisson. 

### Solution
Since Poisson uses averages, and we expect one failure every 10 years, we can say the average yearly failure rate is 0.1.
We're looking for 0 failures in the first 8 years.

$P_{i}(t) = \frac{(\lambda t)^{i}e^{-\lambda t}}{i!}$

$P_{0}(8) = \frac{(0.1\cdot 8)^{0}e^{-0.1\cdot 8}}{0!} = e^{-0.8} =$ `r exp(-0.8)`.

$E[X] = Var(X) = \lambda\cdot t = 0.1 \cdot 8 = 0.8$.

Standard Deviation $= \sqrt{Var(X)} = \sqrt{0.8} =$ `r sqrt(0.8)`.

```{r}
lambda <- 0.1
t <- 8
i <- 0
ex <- lambda*t
var <- lambda*t
sd <- sqrt(var)
pdf <- ppois(i, t*lambda)

cat(sprintf("\n %s = %f \n", 
            c("Probability", "Expected Value", "Variance", "Standard Deviation"),
            c(pdf, ex, var, sd))
    )
```