---
title: "DATA 605 - Assignment 15"
author: "Joshua Sturm"
date: "May 17, 2018"
output:
  pdf_document:
    keep_tex: yes
  html_document:
    highlight: textmate
    theme: sandstone
    code_folding: show
    toc: yes
    toc_float: yes
    smart: yes
  github_document:
always_allow_html: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = F, message = F, collapse = T, cache = T)
setwd("~/GitHub/DATA_605/Assignment 15")
```

```{r}
library(tidyverse)
```


# 1.
Find the equation of the regression line for the given points. Round any final values to the nearest hundredth, if necessary.

\[
( 5.6, 8.8 ), ( 6.3, 12.4 ), ( 7, 14.8 ), ( 7.7, 18.2 ), ( 8.4, 20.8 )
\]

## Solution
\begin{align*}
Y &= mX + b \\
m &= \frac{n\sum(xy) - \sum(x)\sum(y)}{n\sum(x^2) - (\sum x)^2} \\
b &= \frac{\sum(y)\sum(x^2) - \sum(x)\sum(xy)}{n\sum(x^2) - \sum(x)^2}
\end{align*}

```{r}
x <- c(5.6, 6.3, 7, 7.7, 8.4)
y <- c(8.8, 12.4, 14.8, 18.2, 20.8)

lm(y ~ x)
```
\[
Y = -14.8 + 4.26x
\]

```{r}
ggplot(data.frame(x,y), aes(x, y)) +
  geom_point() +
  geom_smooth(method = "lm")
```

# 2.
Find all local maxima, local minima, and saddle points for the function given below. Write your answer(s) in the form
$(x, y, z)$. Separate multiple points with a comma.

\[
f(x, y) = 24x - 6xy^2 - 8y^3
\]

## Solution
We first need to find the first and second partial derivatives.

$f_x = 24 - 6y^2$ \ 
$f_y = -12xy - 24y^2$

$24 - 6y^2 = 0 \ \to \ y^2 = 4 \ \to \ y = \pm 2$

When $y = 2$: \ $-12xy -24y^2 = 0 \ \to \ -24x = 96 \ \to \ x = -4$.

When $y = -2$: \ $-12xy -24y^2 = 0 \ \to \ 24x = 96 \ \to \ x = 4$.

Plugging these values in to get our third coordinate:

$f(-4,2) = 24(-4) - 6(-4)(2^2) - 8(2^3) =$ `r 24*(-4) - 6*(-4)*(2^2) - 8*(2^3)`.

$f(4,-2) = 24(4) - 6(4)(-2^2) - 8(-2^3) =$ `r 24*(4) - 6*(4)*((-2)^2) - 8*((-2)^3)`.

Our two critical points are $(-4,2,-64)$ and $(4,-2,64)$.

To classify these extrema, we can use the second derivative test.

$D(x,y) = f_{xx}f_{yy} - f_{xy}^2$.

$f_{xx} = 0$.

$f_{yy} = -12x - 48y$.

$f_{xy} = f_{yx} = -12y$.

$D = 0 - (-12y)^2 = -144y^2$.

$D(x,y) < 0 \ \forall (x,y)$, so both critical points are saddle points.

# 3.
A grocery store sells two brands of a product, the "house" brand and a "name" brand. The manager estimates that if she
sells the "house" brand for $x$ dollars and the "name" brand for $y$ dollars, she will be able to sell $81 - 21x + 17y$ units of the "house" brand and $40 + 11x - 23y$ units of the "name" brand.
\ 
Step 1. Find the revenue function $R(x, y)$.
\ 
Step 2. What is the revenue if she sells the "house" brand for \$2.30 and the "name" brand for \$4.10?

## Solution

Step 1: The revenue function is simply the number of units sold multipled by the price.

$R(x,y) = (81 - 21x + 17y)x + (40 + 11x - 23y)y$.

Step 2: Plug in $x = 2.30$ and $y = 4.10$.

$R(2.30, 4.10) = \Big[81 - 21(2.30) + 17(4.10)\Big]\times 2.30 + \Big[40 + 11(2.30) - 23(4.10)\Big]\times 4.10$

$R(2.30, 4.10) =$ `r (81 - 21*(2.30) + 17*(4.10))*2.30 + (40 + 11*(2.30) - 23*(4.10))*4.10`. 

# 4.
A company has a plant in Los Angeles and a plant in Denver. The firm is committed to produce a total of 96 units of a
product each week. The total weekly cost is given by $C(x,y) = \frac{1}{6}x^2 + \frac{1}{6}y^2 + 7x + 25y + 700$, where $x$ is the number of units produced in Los Angeles and $y$ is the number of units produced in Denver. How many units should be produced in each plant to minimize the total weekly cost?

## Solution
Total units produced each week = Units produced in Los Angeles + Units produced in Denver.

We're given $x + y = 96$. We need to find the critical points of $C(x,y)$, and then find the local minimum.

Solve for either variable: $x = y - 96, \ \ y = 96 - x$.

\begin{align*}
C(x, 96 - x) &= \frac{1}{6}x^2 + \frac{1}{6}(96 - x)^2 + 7x + 25(96 - x) + 700 \\
&= \frac{1}{6}x^2 + 1536 - 32x + \frac{1}{6}x^2 + 7x + 2400 - 25x + 700 \\
&= \frac{1}{3}x^2 - 50x + 4636 \\
C_x &= \frac{2}{3}x - 50 = 0 \\
x &= 75 \\
C_{xx} &= \frac{2}{3}
\end{align*}

Since the second derivative is > 0, then, by the Second Derivative Test, there is a local minimum at 75.

Thus, the Los Angeles plant should produce 75 units, and the Denver plant should produce 21 units.

# 5.
Evaluate the double integral on the given region. Write your answer in exact form without decimals.

\[
\iint \limits_R (e^{8x + 3y})dA; \ \ R: 2 \leq x \leq 4 \ \ \text{and } \ 2 \leq y \leq 4
\]

## Solution
\begin{align*}
&\int\limits_{2}^{4} \int\limits_{2}^{4} (e^{8x + 3y}) \ dx \ dy \\
&\int\limits_{2}^{4} e^{8x} \ dx \int\limits_{2}^{4} e^{3y} \ dy \\
&\frac{1}{8}e^{8x}\Big|_{2}^{4} \cdot \frac{1}{3}e^{3y}\Big|_{2}^{4} \\
&\frac{1}{24}(e^{32} - e^{16})(e^{12} - e^{6})
\end{align*}

```{r}
(1/24)*((exp(32) - exp(16))*(exp(12) - exp(6)))
```
